### 2024-10-07 Chapters/Chapter3.tex
Final preperations
Updated tables in 5 and added 6
Added abstract. For some reason the citations are not being loaded so just get [descriptive_name].

\begin{figure}[h]
    \caption{A flow chart showing how data sources were combined}
    \label{fig:data_preprocessing_flow_chart}
    \includegraphics[scale = 0.5]{Figures/data-preprocessing-flow-chart.png}
\end{figure}

@@

\begin{table}[h]
    \caption[Comparison of measured and reanalysis wind speed]{Comparison of measured and reanalysis wind speed using mean absolute error (MAE) and mean absolute percentage error (MAPE). Note that for the computation of MAPE for ranges that otherwise include 0, 0 values have been excluded so as to prevent division by zero and exploding values. The comparisons are done using measured wind speed (at 10 meters above ground for IMO and 6-7 meters above ground for IRCA) and reanalysis wind speed at 15 meters above ground.}
    \label{table:measuredVSReanalysis_wind_speed}
    \centering
    \begin{tabular}{c|ccc}
        f & n & MAE & MAPE \\ \hline
        {[}0;5{[} & 6.2e6 & 2.1 & 1.6\\
        {[}5;10{[} & 4.2e6 & 2.2 & 0.3\\
        {[}10;15{[} & 1.5e6 & 2.5 & 0.2\\
        {[}15;20{[} & 3.9e5 & 3.0 & 0.2\\
        {[}20;25{[} & 8.4e4 & 4.0 & 0.2\\
        {[}25;$\infty${[} & 2.0e4 & 6.6 & 0.2\\
        {[}0;$\infty${[} & 1.2e7 & 2.2 & 1.0
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[]{}
    \label{table:station_mae_distribution}
    \resizebox{\textwidth}{!}{
    \centering
    \begin{tabular}{c|ccc}
        Station & Number of measurements & MAE & Location \\\hline
        1470 & 6.8e3 & 1.17 & Reykjavík Háahlíð \\
        1350 & 5.2e4 & 1.18 & Keflavíkurflugvöllur \\
        1482 & 1.4e4 & 1.23 & Reykjavík Víðidalur \\
        4921 & 1.3e4 & 1.29 & Rif á Melrakkasléttu \\
        1477 & 5.6e4 & 1.29 & Reykjavíkurflugvöllur\\ \hdashline[0.5pt/1pt]
        35553 & 4.0e3 & 4.30 & Almannaskarð - göng\\
        6745 & 1.5e4 & 4.36 & Kerlingarfjöll - Ásgarðsfjall\\
        35978 & 7.9e3 & 4.40 & Fáskrúðsfjarðargöng suður\\
        2640 & 1.6e3 & 4.51 & Seljalandsdalur\\
        32635 & 3.2e4 & 4.95 & Botn í Súgandafirði\\
    \end{tabular}
    }
\end{table}

>>>>

\begin{figure}[h]
    \includegraphics[scale = 0.5]{Figures/data-preprocessing-flow-chart.png}
    \caption{A flow chart showing how data sources were combined}
    \label{fig:data_preprocessing_flow_chart}
\end{figure}

@@

\begin{table}[h]
    \caption[Comparison of measured and reanalysis wind speed]{Comparison of measured and reanalysis wind speed using mean absolute error (MAE) and mean absolute percentage error (MAPE). Note that for the computation of MAPE for ranges that otherwise include 0, 0 values have been excluded so as to prevent division by zero and exploding values. The comparisons are done using measured wind speed (at 10 meters above ground for IMO and 6-7 meters above ground for IRCA) and reanalysis wind speed at 15 meters above ground.}
    \label{table:measuredVSReanalysis_wind_speed}
    \centering
    \begin{tabular}{cccc}

        f & n & MAE & MAPE \\

        {[}0;5{[} & 6.2e6 & 2.1 & 1.6\\
        {[}5;10{[} & 4.2e6 & 2.2 & 0.3\\
        {[}10;15{[} & 1.5e6 & 2.5 & 0.2\\
        {[}15;20{[} & 3.9e5 & 3.0 & 0.2\\
        {[}20;25{[} & 8.4e4 & 4.0 & 0.2\\
        {[}25;$\infty${[} & 2.0e4 & 6.6 & 0.2\\
        {[}0;$\infty${[} & 1.2e7 & 2.2 & 1.0\\

    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Mean absolute difference of measured wind speed and reanalysis wind speed at select stations]{Mean absolute error for reanalysis wind speed as compared to measured wind speed, for the five stations with the highest difference and the five stations with the lowest difference.}
    \label{table:station_mae_distribution}
    \resizebox{\textwidth}{!}{
    \centering
    \begin{tabular}{cccc}

        Station & Number of measurements & MAE & Location \\

        1470 & 6.8e3 & 1.17 & Reykjavík Háahlíð \\
        1350 & 5.2e4 & 1.18 & Keflavíkurflugvöllur \\
        1482 & 1.4e4 & 1.23 & Reykjavík Víðidalur \\
        4921 & 1.3e4 & 1.29 & Rif á Melrakkasléttu \\
        1477 & 5.6e4 & 1.29 & Reykjavíkurflugvöllur\\ \hdashline[0.5pt/1pt]
        35553 & 4.0e3 & 4.30 & Almannaskarð - göng\\
        6745 & 1.5e4 & 4.36 & Kerlingarfjöll - Ásgarðsfjall\\
        35978 & 7.9e3 & 4.40 & Fáskrúðsfjarðargöng suður\\
        2640 & 1.6e3 & 4.51 & Seljalandsdalur\\
        32635 & 3.2e4 & 4.95 & Botn í Súgandafirði\\

    \end{tabular}
    }
\end{table}

### 2024-10-07 Chapters/Chapter5.tex
Updated tables in 5 and added 6
Added abstract. For some reason the citations are not being loaded so just get [descriptive_name].

\begin{table}[h]
    \caption[Model results for different AWSL]{MAPE for each average wind speed limit with and without landscape elevation in a 30° sector around the point of interest into the direction of the reanalysis wind. The influence of adding elevation data seems to reduce the error. The percentage error is higher for lower wind speeds and thus observing the error for different lower bound of wind speed will produce different results. This lower bound is determined using the reanalysis wind speed at 15 meters ($ws_{15}$).}
    \label{table:results}
    \centering
    \begin{tabular}{lccc}
        \toprule
        \textbf{AWSL} & \textbf{MAPE} & &\\ 
        $[m/s]$ & \textit{Baseline} &  \textit{Without DEM} & \textit{With DEM} \\\hline
        $\geq 0$ & 39.2\% & 19.7\% & 18.6\% \\
        $\geq 5$ & 28.1\% & 15.9\% & 14.1\%\\
        $\geq 10$ & 23.9\% & 13.6\% & 11.5\%\\
        $\geq 15$ & 23.2\% & 13.0\% & 10.6\%\\
        $\geq 20$ & 24.7\% & 13.6\% & 11.1\%\\
        $\geq 25$ & 27.7\% & 15.7\% & 12.9\%\\
        \bottomrule
    \end{tabular}
\end{table}

%m/s & $ws_{15}$ & f & $ws_{15}$ & f & $ws_{15}$ & f\\\hline
%$\geq 0$ & 39.2\% & 39.2\% & 19.7\% & 15.9\% & 18.6\% & 15.3\%\\
%$\geq 5$ & 28.1\% & 14.8\% & 15.9\% & 9.9\% & 14.1\% & 9.4\%\\
%$\geq 10$ & 23.9\% & 11.1\%& 13.6\% & 7.6\% & 11.5\% & 7.4\%\\
%$\geq 15$ & 23.2\% & 9.3\% & 13.0\% & 7.0\% & 10.6\% & 6.4\%\\
%$\geq 20$ & 24.7\% & 8.2\% & 13.6\% & 6.3\% & 11.1\% & 5.8\%\\
%$\geq 25$ & 27.7\% & 7.3\% & 15.7\% & 5.8\% & 12.9\% & 5.5\% \\
%7.5 & 25.3\% & 12.6\% & 14.5\% & 8.6\% & 12.5\% & 8.6\%\\
%12.5 & 23.3\% & 10.1\% & 13.1\% & 7.1\% & 11.0\% & 6.9\%\\
%17.5 & 23.7\% & 8.7\% & 12.8\% & 6.4\% & 10.5\% & 6.1\%\\
%22.5 & 26.1\% & 7.8\% & 14.5\% & 6.1\% & 12.6\% & 5.7\%\\

@@

\begin{table}[h]
    \caption[Model result by station]{The MAPE results for selected stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the AWSL is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:specific_sites}
    \centering
    \begin{tabular}{lccc}
        \textbf{Station name} & \textbf{Number of measurement points} & \multicolumn{2}{c}{\textbf{MAPE}}\\
        & & General training & Site training\\\hline
        Akrafjall & 42,791 & 18.6\% & 93.7\%\\
        Almannaskarð & 4,014 & 12.2\% & 86.7\%\\
        Ásgarðsfjall & 15,121 & 9.1\% & 9.4\%\\
        Jökulheimar & 17,176 & 7.7\% & 7.7\%\\
        Sandbúðir & 18,718 & 6.8\% & 6.4\%\\
        Stórholt & 35,126 & 7.1\% & 29.2\% \\
        Þúfuver & 19,538 & 6.4\% & 6.8\%\\
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Model result looking at closed wind speed intervals]{The MAPE results for different AWSL intervals. Here instead of training for all data above a certain threshold put of the wind, training is done only on data between two wind speeds. The percentage variance in gust factor as a function of wind speed increases with decreasing wind speed. Measured wind speed is used for the cutoff and thus have data leakage. This results should thus be somewhat comparable to the last column in Table (\ref{table:results})}
    \label{table:closed_intervals}
    \centering
    \begin{tabular}{l | cc}
        \textbf{Interval} &  \multicolumn{2}{c}{\textbf{MAPE}}\\
        $[m/s]$ & Without Elevation & With Elevation\\\hline
        $[5, 10[$ & 17.2\% & 15.4\%\\ %10.6\%\\
        $[10, 15[$ & 14.1\% & 11.9\%\\ %7.8\%\\
        $[15, 20[$ & 13.2\% & 10.9\%\\ %6.4\%\\
        $[20, 25[$ & 14.9 \% & 11.6\%\\ %6.3\%\\
        $[25, 30[$ & 17.0\% & 16.0\%\\ %7.1\%\\
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Model result by stations of interest]{The MAPE results of different stations for several stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the AWSL is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:specific_sites}
    \centering
    \begin{tabular}{l | cc}
        \toprule
        \textbf{Station name} & \multicolumn{2}{c}{\textbf{MAPE}}\\
         & Baseline & Model\\
         \midrule
        Fáskrúðsfjörður & 28.2\% & 21.8\%\\
        Ingólfsfjall & 30.0\% & 19.6\%\\
        Kjalarnes & 20.7\% & 13.5\% \\
        Sandskeið & 13.0\% & 10.2\%\\
        Seyðisfjörður & 32.1\% & 23.0\%\\
        Þjórsárdalur & 12.2\% & 11.4\%\\
        Þrengsli & 13.6\% & 11.3\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

In each of these plots, even without the feature labels, the station elevation is easily noticed as the value is constant for a station. This is not noteworthy. What is noteworthy is the range of impact from this single value. For simpler models, this would not happen. It is important to note that SHAP assumes feature independece\cite{Salih_2024}. This might explain why the impact of the Richardson number is so low. Both the squared Brunt–Väisälä frequency and the Richardson number are derived features from reanalysis data. They carry with them some extra information over the other features in the dataset. This is because both are variables over elevation ranges. That is, as seen in Equations (\ref{eqn:Ri}, \ref{eqn:N}), both are dependent on values at lower and upper elevations and try to describe the stability of that range. Shapley tries to assign contribution values for each feature for each observation. SHAP assumes that the features are independent, but this is not the case. It is clearly not the case for the derived variables, but how the contribution should be distributed between the features is not clear. Seemingly the SHAP python package is giving all the impact to the Brunt–Väisälä squared frequency and none to Richardson number. If the Brunt–Väisälä would be excluded from the data, the impact of the Richardson number would likely increase. Another point to note is that the features are ordered by their impact. This means that the station elevation is the most impactful for each plot, but the ordering of other variables changes. Looking at Figure (\ref{fig:ShapleySummary3}), which shows the Shapley summary plot for all data, the wind speed is the second most important feature. This is reversed in Figure (\ref{fig:ShapleySummaryAkrafjall}). This is not unexpected as Akrafjall station was specifically selected as the MAE for reanalysis wind speed was very high as can be seen in Table (\ref{fig:ShapleySummaryAkrafjall}), where you will also find the stations whose summary plot is shown in Figures (\ref{fig:ShapleySummaryAlmannaskarð}, \ref{fig:ShapleySummaryAsgarðsfjall}) and these also fall into the category of very high MAE for wind speed. What is interesting is that the reanalysis wind speed is also of low impact at stations like Háahlíð and Keflavíkurflugvöllur, as shown in Figures (\ref{fig:ShapleySummaryHaahlid}, \ref{fig:ShapleySummaryKeflavikurflugvollur}). These stations had the lowest of MAE for difference between measured wind speed and reanalysis wind speed. As the summary plot over all stations (Figure (\ref{fig:ShapleySummary3})) shows that reanalysis wind speed is impactful, might lead to the conclusion that the reanalysis wind speed is not a good predictor at these locations or something else is skewing the data.

A simpler way to look at feature importance is to create models that are trained on and use to predict different sets of parameters. The results of such a comparison can be seen in Table \ref{table:setsOfParams}.

\begin{table}[h]
    \caption[Model results for different sets of parameters.]{...}
    \label{table:setsOfParams}
    \centering
    \begin{tabular}{lc}
        \toprule
        \textbf{Model parameters} & \textbf{MAPE}\\ 
        \midrule
        Baseline constant & 23.9\%\\
        $ws_{15}$ & 17.2\%\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}]$ 16.7\% & ---\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, ASL, twd]$ & ---\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, ASL, twd, N, Ri]$ & ---\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, ASL, twd, N, Ri]$ + DEM & ---\\
        $[ws_{15,250,500}, t_{15,250,500}, p_{15,250,500}, wd_{15,250,500}, ASL, twd, N, Ri]$
        \bottomrule
    \end{tabular}
\end{table}

>>>>

\begin{table}[h]
    \caption[Model results for different AWSL]{MAPE for each average wind speed limit with and without landscape elevation in a 30° sector around the point of interest into the direction of the reanalysis wind. The influence of adding elevation data seems to reduce the error. The percentage error is higher for lower wind speeds and thus observing the error for different lower bound of wind speed will produce different results. This lower bound is determined using the reanalysis wind speed at 15 meters ($ws_{15}$).}
    \label{table:results}
    \centering
    \begin{tabular}{lccc}
        \toprule
        \textbf{AWSL} & \textbf{MAPE} & &\\ 
        $[m/s]$ & \textit{Baseline} &  \textit{Without DEM} & \textit{With DEM} \\
        \midrule
        $\geq 0$ & 39.2\% & 19.7\% & 18.6\% \\
        $\geq 5$ & 28.1\% & 15.9\% & 14.1\%\\
        $\geq 10$ & 23.9\% & 13.6\% & 11.5\%\\
        $\geq 15$ & 23.2\% & 13.0\% & 10.6\%\\
        $\geq 20$ & 24.7\% & 13.6\% & 11.1\%\\
        $\geq 25$ & 27.7\% & 15.7\% & 12.9\%\\
        \bottomrule
    \end{tabular}
\end{table}

%m/s & $ws_{15}$ & f & $ws_{15}$ & f & $ws_{15}$ & f\\
%$\geq 0$ & 39.2\% & 39.2\% & 19.7\% & 15.9\% & 18.6\% & 15.3\%\\
%$\geq 5$ & 28.1\% & 14.8\% & 15.9\% & 9.9\% & 14.1\% & 9.4\%\\
%$\geq 10$ & 23.9\% & 11.1\%& 13.6\% & 7.6\% & 11.5\% & 7.4\%\\
%$\geq 15$ & 23.2\% & 9.3\% & 13.0\% & 7.0\% & 10.6\% & 6.4\%\\
%$\geq 20$ & 24.7\% & 8.2\% & 13.6\% & 6.3\% & 11.1\% & 5.8\%\\
%$\geq 25$ & 27.7\% & 7.3\% & 15.7\% & 5.8\% & 12.9\% & 5.5\% \\
%7.5 & 25.3\% & 12.6\% & 14.5\% & 8.6\% & 12.5\% & 8.6\%\\
%12.5 & 23.3\% & 10.1\% & 13.1\% & 7.1\% & 11.0\% & 6.9\%\\
%17.5 & 23.7\% & 8.7\% & 12.8\% & 6.4\% & 10.5\% & 6.1\%\\
%22.5 & 26.1\% & 7.8\% & 14.5\% & 6.1\% & 12.6\% & 5.7\%\\

@@

\begin{table}[h]
    \caption[Model result by station]{The MAPE results for selected stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the AWSL is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:specific_sites}
    \centering
    \begin{tabular}{lccc}
        \toprule
        \textbf{Station name} & \textbf{Number of measurement points} & \multicolumn{2}{c}{\textbf{MAPE}}\\
        & & General training & Site training\\
        \midrule
        Akrafjall & 42,791 & 18.6\% & 93.7\%\\
        Almannaskarð & 4,014 & 12.2\% & 86.7\%\\
        Ásgarðsfjall & 15,121 & 9.1\% & 9.4\%\\
        Jökulheimar & 17,176 & 7.7\% & 7.7\%\\
        Sandbúðir & 18,718 & 6.8\% & 6.4\%\\
        Stórholt & 35,126 & 7.1\% & 29.2\% \\
        Þúfuver & 19,538 & 6.4\% & 6.8\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Model result looking at closed wind speed intervals]{The MAPE results for different AWSL intervals. Here instead of training for all data above a certain threshold put of the wind, training is done only on data between two wind speeds. The percentage variance in gust factor as a function of wind speed increases with decreasing wind speed. Measured wind speed is used for the cutoff and thus have data leakage. This results should thus be somewhat comparable to the last column in Table (\ref{table:results})}
    \label{table:closed_intervals}
    \centering
    \begin{tabular}{lcc}
        \toprule
        \textbf{Interval} &  \multicolumn{2}{c}{\textbf{MAPE}}\\
        $[m/s]$ & Without Elevation & With Elevation\\
        \midrule
        $[5, 10[$ & 17.2\% & 15.4\%\\ %10.6\%\\
        $[10, 15[$ & 14.1\% & 11.9\%\\ %7.8\%\\
        $[15, 20[$ & 13.2\% & 10.9\%\\ %6.4\%\\
        $[20, 25[$ & 14.9 \% & 11.6\%\\ %6.3\%\\
        $[25, 30[$ & 17.0\% & 16.0\%\\ %7.1\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Model result by stations of interest]{The MAPE results of different stations for several stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the AWSL is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:more_specific_sites}
    \centering
    \begin{tabular}{lcc}
        \toprule
        \textbf{Station name} & \multicolumn{2}{c}{\textbf{MAPE}}\\
         & Baseline & Model\\
        \midrule
        Fáskrúðsfjörður & 28.2\% & 21.8\%\\
        Ingólfsfjall & 30.0\% & 19.6\%\\
        Kjalarnes & 20.7\% & 13.5\% \\
        Sandskeið & 13.0\% & 10.2\%\\
        Seyðisfjörður & 32.1\% & 23.0\%\\
        Þjórsárdalur & 12.2\% & 11.4\%\\
        Þrengsli & 13.6\% & 11.3\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

In each of these plots, even without the feature labels, the station elevation is easily noticed as the value is constant for a station. This is not noteworthy. What is noteworthy is the range of impact from this single value. For simpler models, this would not happen. It is important to note that SHAP assumes feature independece\cite{Salih_2024}. This might explain why the impact of the Richardson number is so low. Both the squared Brunt–Väisälä frequency and the Richardson number are derived features from reanalysis data. They carry with them some extra information over the other features in the dataset. This is because both are variables over elevation ranges. That is, as seen in Equations (\ref{eqn:Ri}, \ref{eqn:N}), both are dependent on values at lower and upper elevations and try to describe the stability of that range. Shapley tries to assign contribution values for each feature for each observation. SHAP assumes that the features are independent, but this is not the case. It is clearly not the case for the derived variables, but how the contribution should be distributed between the features is not clear. Seemingly the SHAP python package is giving all the impact to the Brunt–Väisälä squared frequency and none to Richardson number. If the Brunt–Väisälä would be excluded from the data, the impact of the Richardson number would likely increase. Another point to note is that the features are ordered by their impact. This means that the station elevation is the most impactful for each plot, but the ordering of other variables changes. Looking at Figure (\ref{fig:ShapleySummary3}), which shows the Shapley summary plot for all data, the wind speed is the second most important feature. This is reversed in Figure (\ref{fig:ShapleySummaryAkrafjall}). This is not unexpected as Akrafjall station was specifically selected as the MAE for reanalysis wind speed was very high as can be seen in Table (\ref{fig:ShapleySummaryAkrafjall}), where you will also find the stations whose summary plot is shown in Figures (\ref{fig:ShapleySummaryAlmannaskarð}, \ref{fig:ShapleySummaryAsgarðsfjall}) and these also fall into the category of very high MAE for wind speed. What is interesting is that the reanalysis wind speed is also of low impact at stations like Háahlíð and Keflavíkurflugvöllur, as shown in Figures (\ref{fig:ShapleySummaryHaahlid}, \ref{fig:ShapleySummaryKeflavikurflugvollur}). These stations had the lowest of MAE for difference between measured wind speed and reanalysis wind speed. As the summary plot over all stations (Figure (\ref{fig:ShapleySummary3})) shows that reanalysis wind speed is impactful, might lead to the conclusion that the reanalysis wind speed is not a good predictor at these locations or something else is skewing the data. A simpler way to look at feature importance is to create models that are trained on and use to predict different sets of parameters. The results of such a comparison can be seen in Table \ref{table:setsOfParams}.

\begin{table}[h]
    \caption[Model results for different sets of parameters.]{...}
    \label{table:setsOfParams}
    \centering
    \begin{tabular}{lc}
        \toprule
        \textbf{Model parameters} & \textbf{MAPE}\\
        \midrule
        Baseline constant & 23.9\%\\
        $ws_{15}$ & 17.2\%\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}]$ & 16.7\% \\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, ASL, twd]$ & 13.8\% \\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, ASL, twd, N, Ri]$ & ---\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, ASL, twd, N, Ri]$ + DEM & ---\\
        $[ws_{15,250,500}, t_{15,250,500}, p_{15,250,500}, wd_{15,250,500}, ASL, twd, N, Ri]$ & --- \\
        \bottomrule
    \end{tabular}
\end{table}

### 2024-10-07 MSc.tex
Added abstract. For some reason the citations are not being loaded so just get [descriptive_name].

\setkomafont{section}{\huge} % The title "Abstract" and "Útfdráttur" should look like the chapters, i.e. use \huge (\chapter cannot be used as this would create a new page)
\section*{Abstract}
In many industries, whether it be windmill farming or transportation, being able to precisely predict the load caused by weather is crucial to prevent harm coming to people and machinery as well as increasing productivity. The efficiency of wind turbines in low wind speed conditions is a critical challenge in optimizing renewable energy systems. Gust predictions can be instrumental in loss prevention. If looking only at wind, the most destructive moments are gusts, by definition. To predict these values there are two ways, traditional numerical weather prediction systems and machine learning or other pattern recognition systems. Computer aided numerical weather prediction systems originated in the 1950's, while the use of machine learning in numerical weather predictions is much more recent, having truly gained traction in the last decade.

The goal of this thesis is try to combine these two methods to try and predict wind gust factor, where the gust factor is defined as the ratio between wind gust and average wind speed. This is done by having the input data to the machine learning model be reanalysis variables that were generated by traditional numerical weather prediction systems. In addition to these variables and derived variables, a digital elevation model (DEM) data was used. A deep neural network was constructed using some set of variables and the results are compared to a baseline as well as each other so as to show the impact each feature has.
\vfill \vspace*{1cm}

\section*{Útdráttur}
Í mörgum iðnuðum, hvort sem það er vindmyllubúgarður eða flutningar, að geta 
\vfill

>>>>

\setkomafont{section}{\huge} % The title "Abstract" and "Útdráttur" should look like the chapters, i.e. use \huge (\chapter cannot be used as this would create a new page)

\section*{Abstract}
In many industries, whether it be windmill farming or transportation, being able to precisely predict the load caused by weather is crucial to prevent harm coming to people and machinery as well as increasing productivity. The efficiency of wind turbines in high wind speed conditions is a vital challenge in optimizing renewable energy systems. Gust predictions can be instrumental in loss prevention. If looking only at wind, the most destructive moments are gusts, by definition. To predict these values there are two ways, traditional numerical weather prediction systems and machine learning or other pattern recognition systems. Computer aided numerical weather prediction systems originated in the 1950's, while the use of machine learning in numerical weather predictions is much more recent, having truly gained traction in the last decade. The goal of this thesis is try to combine these two methods to try and predict wind gust factor, where the gust factor is defined as the ratio between wind gust and average wind speed. This is done by having the input data to the machine learning model be reanalysis variables that were generated by traditional numerical weather prediction systems. In addition to these variables and derived variables, a digital elevation model (DEM) data was used. A deep neural network was constructed using some set of variables and the results are compared to a baseline as well as each other so as to show the impact each feature has.
\vfill \vspace*{1cm}

\section*{Útdráttur}
Í mörgum iðnuðum, hvort sem það er vindmyllubúgarður eða flutningar er getan til þess að spá fyrir um álag vegna veðurs mikilvæg til að koma í veg fyrir tjóns á vélum og byggingum ásamt því að koma í veg fyrir skaða á fólki. Einnig er hægt að auka framleiðni. Nýtni vindtúrbínna í hávinda aðstæðum er mikilvæg áskorun í bestun endurnýjanlega orkukerfisins. Hviðuspár geta verið stór þáttur í tjónaminnkun. Ef einungis er horft á vind, þá eru hviður hæsti álagspunktur út frá skilgreiningu. Sögulega hefur verið notast við hefðbundin spálíkön sem reiða sig á töluleg eðlisfræðileg líkön til að spá fyrir um niðurstöður. Á síðustu árum hefur gervigreind þróast mikið og líkön verið þjálfuð til þess að giska á veður. Svo miklar framfarir hafa orðið að sum líkön geta keppt við hefðbundin veðurlíkön. Markmið þessa verkefnis er reyna að nýta bæði hefðbundin veðurlíkön og gervigreind til þess að giska á hviðustuðla, þar sem hviðustuðullinn er skilgreindur sem hlutfall hæstu hviðu og meðalvinds. Þetta er gert með því að nota endurgreiningargögn úr hefðbundnu veðurlíkani sem grunngögn í gervigreindarlíkan. Einnig eru notað hæðarlíkan til að lýsa umhverfi. Djúpt tauganet var búið til með því að velja breytur úr endurgreiningargögnunum ásamt hæðarpunktum og afleiddum breytum. Þetta líkan er svo þjálfað og niðurstöður bornar saman við grunnlíkan sem og önnur líkön með öðrum breytum til að skoða áhrif breyta.
\vfill
